---
title: "**Appendix S2 - Analysis of adult giant anteater movement**"
author: "Aimee Chhen, Alessandra Bertassoni, Arnaud LJ Desbiez, Michael J Noonan"
date: "This document was created on `r format(Sys.Date(), '%B %d, %Y')`."
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, warning = FALSE, message = FALSE}
# Load packages
library(ctmm)
library(dplyr)
library(tidyverse)
library(glmmTMB)
library(lme4)

```

The pre-print analyses contained 23 range-resident giant anteaters. Within the 23 individuals contained 12 males (8 adults, 4 subadults) and 11 females (all adults).
This script is based on adult giant anteaters only. Subadults have been removed.

# Data

```{r import-data, eval = FALSE, message = FALSE}
#import data, cleaned giant anteater GPS tracking data, containing no outliers
COLLAR_DATA <- read.csv("data/anteater/Anteaters_NoOutliers.csv")

#correct mismatch ID entries
COLLAR_DATA$ID[COLLAR_DATA$ID == "Larry 267"] <- "Larry"

COLLAR_DAT <- data.frame(timestamp = COLLAR_DATA$timestamp,
                         ID = COLLAR_DATA$ID,
                         GPS.Latitude = COLLAR_DATA$GPS.Latitude,
                         GPS.Longitude = COLLAR_DATA$GPS.Longitude,
                         GPS.Horizontal.Dilution = COLLAR_DATA$GPS.Horizontal.Dilution,
                         OUT = COLLAR_DATA$OUT,
                         Road = COLLAR_DATA$Road)

#subset to the 20 adult range-resident individuals
anteater_data <- COLLAR_DAT[which(COLLAR_DAT$ID 
                                  %in% c("Alexander", "Annie", "Beto", "Bumpus", "Cate", 
                                         "Christoffer","Elaine", "Hannah","Jackson", "Jane", 
                                         "Larry", "Luigi", "Makao", "Margaret", "Maria", 
                                         "Puji", "Rodolfo", "Sheron", "Thomas")),]

save(anteater_data, file = "data/anteater/anteater_data_adults.rda")
```

```{r}
load("data/anteater/anteater_data_adults.rda")
```

```{r message =FALSE}
#import supplementary data containing biological information
DATA_META <- read.csv("data/anteater/Anteater_Results_Final.csv")

#subset biological data from supplementary data
bio_data <- DATA_META[which(DATA_META$ID 
                            %in% c("Alexander", "Annie", "Beto", "Bumpus", "Cate", 
                                   "Christoffer","Elaine", "Hannah","Jackson", "Jane", 
                                   "Larry", "Luigi", "Makao", "Margaret", "Maria", 
                                   "Puji", "Rodolfo", "Sheron", "Thomas")),c(1:5)]

#add site location 
bio_data$Site[bio_data$Road == "MS-040"] <- 1
bio_data$Site[bio_data$Road == "BR_267"] <- 2

#add bio data
anteater_data <- merge(anteater_data, bio_data, all = TRUE)

#convert dataset to a telemetry object
DATA_TELEMETRY <- as.telemetry(anteater_data)

```

```{r}
#import movement models
load("data/anteater_fit.rda")

#subset to adults only
FIT <- FIT[-c(3,12,14,20)]

#import home-range estimates
load("data/anteater_akdes.rda")

#subset to adults only
AKDE <- AKDE[-c(3,12,14,20)]

```

# Home-range results

```{r}
#Mean HR size...
meta(AKDE)

#Compare between sex...
AKDE_male <- AKDE[c("Alexander", "Beto","Christoffer","Jackson", "Larry", 
                    "Luigi", "Rodolfo", "Thomas")]
AKDE_female <- AKDE[c("Annie", "Bumpus", "Cate", "Elaine", "Hannah",
                      "Jane","Makao", "Margaret", "Maria", "Puji",
                      "Sheron")]

#calculate mean home range sizes for male
meta(AKDE_male)

#calculate mean home range sizes for female
meta(AKDE_female)

#female/male ratio of mean home-range areas
AKDE_sex_compare <- list(male = AKDE_male,
                         female = AKDE_female)
meta(AKDE_sex_compare, col = c("#004488", "#A50026"), sort = TRUE)

```


# Estimating home-range overlap

```{r}
#subset home range overlap based on site location
AKDE_1 <- AKDE[c("Alexander", "Bumpus", "Cate", "Christoffer",
                 "Elaine", "Jackson", "Makao",
                 "Puji", "Rodolfo")]

AKDE_2 <- AKDE[c("Annie", "Beto", "Hannah", "Jane", "Larry",
                 "Luigi", "Margaret", "Maria", "Sheron",
                 "Thomas")]

#calculate 95% AKDE home range overlap for a pairwise comparison for each site
overlap1 <- overlap(AKDE_1, level = 0.95)
overlap2 <- overlap(AKDE_2, level = 0.95)
```


```{r eval=FALSE}
#indicate the matrix layers to be extracted
matrix_level <- c('low', 'est', 'high')

num_matrices1 <- dim(overlap1$CI)[3]
overlap_result1 <- list()
for (i in 1:num_matrices1) {
  matrix_layer1 <- overlap1$CI[,,i]
  matrix_layer1[upper.tri(matrix_layer1, diag = TRUE)] <- NA
  matrix_layer1 <- as.data.frame(matrix_layer1)
  matrix_layer1$anteater_A <- rownames(matrix_layer1)
  overlap_col_name1 <- paste0('overlap_', matrix_level[i])
  matrix_layer1 <- pivot_longer(matrix_layer1, cols = -anteater_A,
                                names_to = 'anteater_B', values_to = overlap_col_name1,
                                values_drop_na = TRUE)
  overlap_result1[[i]] <- matrix_layer1
}

overlap_df1 <- overlap_result1[[1]]
for (i in 2:num_matrices1) {
  overlap_df1 <- left_join(overlap_df1, overlap_result1[[i]],
                           by = c("anteater_A", "anteater_B"))
}


# Add biological data to data frame
overlap_df1 <- left_join(overlap_df1, rename(bio_data, anteater_A = ID), by = "anteater_A")
overlap_df1 <- left_join(overlap_df1, rename(bio_data, anteater_B = ID), by = "anteater_B", suffix = c(".A", ".B"))

# Add column to indicate which sexes are being compared
overlap_df1 <- mutate(
  overlap_df1, sex_comparison =
    case_when(paste(Sex.A, Sex.B) == "Male Male" ~ "male-male",
              paste(Sex.A, Sex.B) == "Female Female" ~ "female-female",
              paste(Sex.A, Sex.B) == "Male Female" ~ "female-male",
              paste(Sex.A, Sex.B) == "Female Male" ~ "female-male"))

num_matrices2 <- dim(overlap2$CI)[3]
overlap_result2 <- list()
for (i in 1:num_matrices2) {
  matrix_layer2 <- overlap2$CI[,,i]
  matrix_layer2[upper.tri(matrix_layer2, diag = TRUE)] <- NA
  matrix_layer2 <- as.data.frame(matrix_layer2)
  matrix_layer2$anteater_A <- rownames(matrix_layer2)
  overlap_col_name2 <- paste0('overlap_', matrix_level[i])
  matrix_layer2 <- pivot_longer(matrix_layer2, cols = -anteater_A,
                                names_to = 'anteater_B', values_to = overlap_col_name2,
                                values_drop_na = TRUE)
  overlap_result2[[i]] <- matrix_layer2
}

overlap_df2 <- overlap_result2[[1]]
for (i in 2:num_matrices2) {
  overlap_df2 <- left_join(overlap_df2, overlap_result2[[i]],
                           by = c("anteater_A", "anteater_B"))
}

# Add biological data to data frame
overlap_df2 <- left_join(overlap_df2, rename(bio_data, anteater_A = ID), by = "anteater_A")
overlap_df2 <- left_join(overlap_df2, rename(bio_data, anteater_B = ID), by = "anteater_B", suffix = c(".A", ".B"))

# Add column to indicate which sexes are being compared
overlap_df2 <- mutate(
  overlap_df2, sex_comparison =
    case_when(paste(Sex.A, Sex.B) == "Male Male" ~ "male-male",
              paste(Sex.A, Sex.B) == "Female Female" ~ "female-female",
              paste(Sex.A, Sex.B) == "Male Female" ~ "female-male",
              paste(Sex.A, Sex.B) == "Female Male" ~ "female-male"))

#combine both sites into one dataframe
overlap_data <- rbind(overlap_df1, overlap_df2)
overlap_data$pair_ID <- paste(overlap_data$anteater_A, overlap_data$anteater_B, sep = "_")
overlap_data <- relocate(overlap_data, pair_ID, .before = anteater_A)
overlap_data <- relocate(overlap_data, c(overlap_low, overlap_est, overlap_high), .after = sex_comparison)

#clean up environment
rm(matrix_layer1, matrix_layer2, matrix_level, 
   overlap_col_name1, overlap_col_name2,
   overlap_result1, overlap_result2)

overlap_data$site <- overlap_data$Site.A
# save(overlap_data, file = "data/home_range/overlap_data_adult.rda")
```

```{r}
load("data/home_range/overlap_data_adult.rda")
```


# Home-range overlap results

```{r}
#Is sex a factor?...

#rescale the values
min_val <- min(overlap_data$overlap_est)
max_val <- max(overlap_data$overlap_est)
squeeze_min <- 0.001
squeeze_max <- 0.999
overlap_data$overlap_est_squeezed <- ((overlap_data$overlap_est - min_val) / (max_val - min_val)) * (squeeze_max - squeeze_min) + squeeze_min

#compare model with and without sex as a variable
HRO_test <- glmmTMB(overlap_est_squeezed ~ sex_comparison + (1|site), 
                    family = beta_family(link = "logit"), data = overlap_data)
HRO_test2 <- glmmTMB(overlap_est_squeezed ~ 1 + (1|site), 
                     family = beta_family(link = "logit"), data = overlap_data)
HRO_test_results <- anova(HRO_test, HRO_test2)
HRO_test_results

#calculate sex impact via likelihood ratio test
HRO_test_pvalue <- round(HRO_test_results$`Pr(>Chisq)`[2], 2)
HRO_test_pvalue


```


```{r}
#Home range overlap results based on sex comparisons (i.e.. male-male, female-female, and female-male).

#number of home range overlap in each sex comparison category
table(overlap_data$sex_comparison)

#calculate mean home range overlap and the range based on sex comparisons
round(mean(overlap_data$overlap_est[overlap_data$sex_comparison == "male-male"]), 2)
round(min(overlap_data$overlap_est[overlap_data$sex_comparison == "male-male"]), 2)
round(max(overlap_data$overlap_est[overlap_data$sex_comparison == "male-male"]), 2)

round(mean(overlap_data$overlap_est[overlap_data$sex_comparison == "female-female"]), 2)
round(min(overlap_data$overlap_est[overlap_data$sex_comparison == "female-female"]), 2)
round(max(overlap_data$overlap_est[overlap_data$sex_comparison == "female-female"]), 2)

round(mean(overlap_data$overlap_est[overlap_data$sex_comparison == "female-male"]), 2)
round(min(overlap_data$overlap_est[overlap_data$sex_comparison == "female-male"]), 2)
round(max(overlap_data$overlap_est[overlap_data$sex_comparison == "female-male"]), 2)
```


# Estimating proximity ratio



```{r}
#import proximity data
load("data/encounter/proximity_df.rda")
proximity_df <- as.data.frame(proximity_df)

#remove subadults from proximity data
proximity_df <- proximity_df[!(proximity_df$anteater_A %in% "Anthony" | proximity_df$anteater_B %in% "Anthony"),]
proximity_df <- proximity_df[!(proximity_df$anteater_A %in% "Kyle" | proximity_df$anteater_B %in% "Kyle"),]
proximity_df <- proximity_df[!(proximity_df$anteater_A %in% "Little_Rick" | proximity_df$anteater_B %in% "Little_Rick"),]
proximity_df <- proximity_df[!(proximity_df$anteater_A %in% "Reid" | proximity_df$anteater_B %in% "Reid"),]

#combine overlap and proximity data
overlap_data <- merge(overlap_data, proximity_df, all = TRUE)
```

# Proximity ratio results

```{r}
#test for significance in sex, compare model with and without sex as a variable
proximity_test <- glmer(proximity_est ~ sex_comparison + (1|site), 
                        family = Gamma(link = "log"), data = overlap_data)
proximity_test2 <- glmer(proximity_est ~ 1 + (1|site), 
                         family = Gamma(link = "log"), data = overlap_data)
proximity_test_results <- anova(proximity_test, proximity_test2)

#calculate sex impact on proximity ratio via likelihood ratio test
proximity_test_pvalue <- round(proximity_test_results$`Pr(>Chisq)`[2], 2)
proximity_test_pvalue

#compare model with and without overlap as a variable
prox_overlap_test <- glmer(proximity_est ~ overlap_est + (1|site), 
                           family = Gamma(link = "log"), data = overlap_data)
prox_overlap_test2 <- glmer(proximity_est ~ 1 + (1|site), 
                            family = Gamma(link = "log"), data = overlap_data)
prox_overlap_test_results <- anova(prox_overlap_test, prox_overlap_test2)

#calculate home-range overlap impact via likelihood ratio test
prox_overlap_test_pvalue <- round(prox_overlap_test_results$`Pr(>Chisq)`[2], 2)
prox_overlap_test_pvalue
```


# Estimating encounters

```{r}
#import the distance data
DATA_DISTANCE <- readRDS("data/encounter/DATA_DISTANCE.rds")

#locate NA values within the dataframe
DATA_DISTANCE[!complete.cases(DATA_DISTANCE), ] #3,502,701 observations
#drop the 3 fixes that had no distance values 
DATA_DISTANCE <- na.omit(DATA_DISTANCE) #3,502,698 observations

#add overlap and proximity information to the distance dataframe
distance_data <- merge(DATA_DISTANCE, proximity_df, by = "pair_ID")
distance_data <- relocate(distance_data, c(distance_low, distance_est, distance_high,
                                       t, timestamp), .after = proximity_high)

#save the distance dataframe
save(distance_data, file = "data/encounter/distance_data_adults.rda")
```



```{r}
#import distance data
load("data/encounter/distance_data_adults.rda")

overlap_data$encounter_count <- NA
pair_ID <- unique(overlap_data$pair_ID)

#calculate total encounters of all individuals based on sex comparison type
for (i in pair_ID){
  subset_A <- distance_data[distance_data$pair_ID == i,]
  
  #count the number of times distance is below 15
  encounter_count <- sum(subset_A$distance_est < 15)
  
  #save results
  overlap_data[overlap_data$pair_ID == i, "encounter_count"] <- encounter_count
  
}

```

# Encounter results


```{r}

#calculate the number of encounters based on threshold of 15m
sum(overlap_data$encounter_count)
sum(overlap_data$encounter_count[overlap_data$sex_comparison == "male-male"])
sum(overlap_data$encounter_count[overlap_data$sex_comparison == "female-female"])
sum(overlap_data$encounter_count[overlap_data$sex_comparison == "female-male"])

#effect of sex and overlap on encounter rates and does not include 0 encounter counts
encounter_test <- glmer(encounter_count ~ overlap_est + sex_comparison + (1|site), 
                        family = poisson(link = "log"), 
                        data = overlap_data, subset = encounter_count > 0)
encounter_test2 <- glmer(encounter_count ~ 1 + (1|site), 
                         family = poisson(link = "log"), 
                         data = overlap_data, subset = encounter_count > 0)
encounter_test_results <- anova(encounter_test, encounter_test2)
encounter_test_results

#calculate sex impact on encounters via likelihood ratio test
encounter_test_pvalue <- round(encounter_test_results$`Pr(>Chisq)`[2], 2)
encounter_test_pvalue

#amount of home range overlap and the number of observed encounters
summary(encounter_test)
```


# Deviated pairs

```{r}

#identify pairs that did not have a proximity ratio of 1
proximity_above1 <- proximity_df[proximity_df$proximity_low > 1,]
proximity_below1 <- proximity_df[proximity_df$proximity_high < 1,]

#exclude pairs with a home range overlap of 0
proximity_below1[proximity_below1$overlap_est > 0.0001,]
proximity_below1 <- proximity_below1[proximity_below1$overlap_est > 0.0001,]

#create a dataframe of the deviated pairs
proximity_identified_pairs_df <- rbind(proximity_above1, proximity_below1)
proximity_identified_pairs_df$pair_ID_number <- seq(from = 1, to = 7, by = 1)
proximity_identified_pairs_df <- relocate(proximity_identified_pairs_df, pair_ID_number,
                                          .before = anteater_A)

#correct the sex_comparison output to female-male
proximity_identified_pairs_df <- mutate(proximity_identified_pairs_df, sex_comparison = 
                                          case_when(paste(Sex.A, Sex.B) == "Male Male" ~ "male-male",
                                                    paste(Sex.A, Sex.B) == "Female Female" ~ "female-female",
                                                    paste(Sex.A, Sex.B) == "Male Female" ~ "female-male",
                                                    paste(Sex.A, Sex.B) == "Female Male" ~ "female-male"))

#clean up environment
rm(proximity_above1, proximity_below1)

save(proximity_identified_pairs_df, file = "data/encounter/proximity_identified_adult_pairs_df.rda")
```



```{r}
#import identified deviated pairs (7 pairs)
load("data/encounter/proximity_identified_adult_pairs_df.rda")

#number of pairs with a deviated proximity ratio based on sex comparison
table(proximity_identified_pairs_df$sex_comparison)

#test for significance in sex, compare model with and without sex as a variable
proximity_test_pairs <- glmer(proximity_est ~ sex_comparison + (1|site), 
                              family = Gamma(link = "log"), 
                              data = proximity_identified_pairs_df)
proximity_test2_pairs <- glmer(proximity_est ~ 1 + (1|site), 
                               family = Gamma(link = "log"), 
                               data = proximity_identified_pairs_df)
proximity_test_results_pairs <- anova(proximity_test_pairs, proximity_test2_pairs)
proximity_test_results_pairs

#calculate sex impact via likelihood ratio test
proximity_test_pvalue_pairs <- round(proximity_test_results_pairs$`Pr(>Chisq)`[2], 2)
proximity_test_pvalue_pairs

#test for significance in home range overlap, compare model with and without overlap as a variable
prox_overlap_test_pairs <- glmer(proximity_est ~ overlap_est + (1|site), 
                                 family = Gamma(link = "log"), 
                                 data = proximity_identified_pairs_df)
prox_overlap_test2_pairs <- glmer(proximity_est ~ 1 + (1|site), 
                                  family = Gamma(link = "log"), 
                                  data = proximity_identified_pairs_df)
prox_overlap_test_results_pairs <- anova(prox_overlap_test_pairs, prox_overlap_test2_pairs)
prox_overlap_test_results_pairs

#calculate home-range overlap impact via likelihood ratio test
prox_overlap_test_pvalue_pairs <- round(prox_overlap_test_results_pairs$`Pr(>Chisq)`[2], 2)
prox_overlap_test_pvalue_pairs
```



# Investigating if weight is a factor in home-range size

```{r}

#import supplementary data containing biological information
DATA_META <- read.csv("data/anteater/Anteater_Results_Final.csv")
#subset biological data from supplementary data
bio_data <- DATA_META[c(1:3,8:10,12,14,17,19,20,22,23,25:29,33:35,37,38), c(1:5)]
#add site location 
bio_data$Site[bio_data$Road == "MS-040"] <- 1
bio_data$Site[bio_data$Road == "BR_267"] <- 2

#load home range size data
load("data/home_range/HR_size.rda")
Weight <- bio_data[,4]
HR_size <- cbind(HR_size, Weight)

#without luigi
HR_size <- HR_size[HR_size$ID != "Luigi",]

#............................................................
# Is weight a factor in home-range size?

#compare model with and without weight as a variable
HR_weight_test <- glmmTMB(HR_est ~ Weight + (1|Site), 
                          family = Gamma(link = "log"), data = HR_size)
HR_weight_test2 <- glmmTMB(HR_est ~ 1 + (1|Site), 
                           family = Gamma(link = "log"), data = HR_size)

HR_weight_test_results <- anova(HR_weight_test, HR_weight_test2)
HR_weight_test_results

#calculate weight impact via likelihood ratio test
HR_weight_test_pvalue <- round(HR_weight_test_results$`Pr(>Chisq)`[2], 2)
HR_weight_test_pvalue

summary(HR_weight_test)

plot(HR_size$HR_est ~ HR_size$Weight)

```









